PaperOutline_ACEsCog

### 2022-02-21 PaperOutline_ACEsCog.Rmd
item-level analysis of deprivation
deprivation factor score broken down into SES variables and neglect variables

@@

@@

>>>>

## Try a different calculation for "parental monitoring" to try to match ABCD (note: this is really a conceptual stretch...):
```{r}
all.data.complete.ageres$monitoring <- rowMeans(data.frame(as.numeric(all.data.complete.ageres$fes18),as.numeric(all.data.complete.ageres$fes48)),na.rm=TRUE)

# Verify that correlation is low: it's 0.247
cor.test(all.data.complete.ageres$monitoring,all.data.complete.ageres$emo_neglect)

# Test specificity
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+monitoring,data=all.data.complete.ageres[all.data.complete.ageres$selection==1,])) # just monitoring
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+monitoring,data=all.data.complete.ageres[all.data.complete.ageres$selection==1,])) # just monitoring
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+monitoring,data=all.data.complete.ageres[all.data.complete.ageres$selection==1,])) # monitoring .
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+monitoring,data=all.data.complete.ageres[all.data.complete.ageres$selection==1,])) # monitoring .

plot(all.data.complete.ageres[all.data.complete.ageres$selection==1,]$monitoring,all.data.complete.ageres[all.data.complete.ageres$selection==1,]$EF)
```

# Try using just the physical neglect part of the deprivation score rather than the whole thing (which includes SES)
# From Mauricio from Julia: deprivation =~ minstlevel_4cat + abepstrat_r.w0 + pphyneg_r.w0 + cphysneg_r.w0 + faincome_qr.w0 + fstatus_r.w0
```{r}
# merge in new items (NOTE: SUBJECT COUNT DROPS TO N=914)
deprivation_subitems <- readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/DeprivationScores_BHRC/jschafer_p2_prodata_20211003.rds")
colnames(deprivation_subitems)[1]<-"SUBJECTID"
all.data.ageres.dep <- merge(all.data.complete.ageres,deprivation_subitems[,c("SUBJECTID", setdiff(colnames(deprivation_subitems),colnames(all.data.complete.ageres)))],by="SUBJECTID")

# child report of physical neglect score: all.data.ageres.dep$cphysneg_r.w0

# Verify that correlation is low: it's 0.015
cor.test(all.data.ageres.dep$cphysneg_r.w0,all.data.ageres.dep$emo_neglect)

# Test specificity
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+cphysneg_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+cphysneg_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+cphysneg_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+cphysneg_r.w0,data=all.data.ageres.dep)) # neither

# So it's not really due to the neglect variable here, what about ses?
summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+fstatus_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+fstatus_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+fstatus_r.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+fstatus_r.w0,data=all.data.ageres.dep)) # neither

summary(lmer(EF~age+gender+(1|state)+(1|schoolID)+emo_neglect+faincome_qr.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(WM~age+gender+(1|state)+(1|schoolID)+emo_neglect+faincome_qr.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(TP~age+gender+(1|state)+(1|schoolID)+emo_neglect+faincome_qr.w0,data=all.data.ageres.dep)) # just emo neglect
summary(lmer(IC~age+gender+(1|state)+(1|schoolID)+emo_neglect+faincome_qr.w0,data=all.data.ageres.dep)) # neither
```

@@

# make sure monitoring and acceptance are on the same scale so we can compare effect sizes:
abcd.data$acceptance<-(((abcd.data$acceptance-1)/2)*4)+1

@@

### Is parental monitoring related to increased stress exposures? Nope!
```{r}
ksads.data<-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/Strife/abcd_ptsd01.txt",header=TRUE)
ksads.baseline<-ksads.data[ksads.data$eventname=="baseline_year_1_arm_1",]
abcd.data.with.ksads <- merge(abcd.data,ksads.baseline[, c("subjectkey", setdiff(colnames(ksads.baseline),colnames(abcd.data)))],by="subjectkey")

abcd.data.with.ksads$totalACEs <- rowSums(data.frame(as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_754_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_755_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_756_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_757_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_758_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_759_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_760_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_761_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_762_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_763_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_764_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_765_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_766_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_767_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_768_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_769_p),as.numeric(abcd.data.with.ksads$ksads_ptsd_raw_770_p)),na.rm=TRUE)

summary(lmer(totalACEs~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+monitoring,data=abcd.data.with.ksads))
```

# Adding adoption as a covariate 
```{r}
adoption.int.ext<-readRDS("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/DEAP-adoption-internalizing-externalizing.rds")
adoption.int.ext<-adoption.int.ext[adoption.int.ext$event_name=="baseline_year_1_arm_1",]
abcd.data.with.intext <- merge(abcd.data,adoption.int.ext[, c("src_subject_id", setdiff(colnames(adoption.int.ext),colnames(abcd.data)))],by="src_subject_id")

abcd.data.with.intext$adopted <- ifelse(!is.na(abcd.data.with.intext$demo_adopt_age_p),1,0)

summary(lmer(thompson_PC1~interview_age+sex+adopted+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.with.intext))
summary(lmer(thompson_PC2~interview_age+sex+adopted+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.with.intext))
summary(lmer(thompson_PC3~interview_age+sex+adopted+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.with.intext))

summary(lmer(acceptance~interview_age+sex+adopted+(1|rel_family_id)+(1|abcd_site),data=abcd.data.with.intext))
summary(lmer(monitoring~interview_age+sex+adopted+(1|rel_family_id)+(1|abcd_site),data=abcd.data.with.intext))

## looking for double dissociation with internalizing/externalizing symptoms (it's not really there)
```{r}
summary(lmer(cbcl_scr_syn_internal_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.with.intext))
summary(lmer(cbcl_scr_syn_external_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.with.intext))

summary(lmer(cbcl_scr_syn_internal_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance,data=abcd.data.with.intext))
summary(lmer(cbcl_scr_syn_internal_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+monitoring,data=abcd.data.with.intext))
summary(lmer(cbcl_scr_syn_external_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance,data=abcd.data.with.intext))
summary(lmer(cbcl_scr_syn_external_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+monitoring,data=abcd.data.with.intext))

# what about other measures from CBCL?
cbcls <-read.table("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/aseba/abcd_cbcls01.txt",header=TRUE)
cbcls.baseline<-cbcls[cbcls$eventname=="baseline_year_1_arm_1",]
abcd.data.cbcls <- merge(abcd.data,cbcls.baseline[, c("subjectkey", setdiff(colnames(cbcls.baseline),colnames(abcd.data)))],by="subjectkey")
abcd.data.cbcls$cbcl_scr_syn_anxdep_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_anxdep_r)
abcd.data.cbcls$cbcl_scr_syn_aggressive_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_aggressive_r)
abcd.data.cbcls$cbcl_scr_syn_social_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_social_r)
abcd.data.cbcls$cbcl_scr_syn_thought_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_thought_r)
abcd.data.cbcls$cbcl_scr_syn_somatic_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_somatic_r)
abcd.data.cbcls$cbcl_scr_syn_withdep_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_withdep_r)
abcd.data.cbcls$cbcl_scr_syn_rulebreak_r<-as.numeric(abcd.data.cbcls$cbcl_scr_syn_rulebreak_r)

summary(lmer(cbcl_scr_syn_aggressive_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.cbcls))

abcd.data.cbcls$cbcl_scr_dsm5_depress_r<-as.numeric(abcd.data.cbcls$cbcl_scr_dsm5_depress_r)
abcd.data.cbcls$cbcl_scr_dsm5_anxdisord_r<-as.numeric(abcd.data.cbcls$cbcl_scr_dsm5_anxdisord_r)
abcd.data.cbcls$cbcl_scr_dsm5_adhd_r<-as.numeric(abcd.data.cbcls$cbcl_scr_dsm5_adhd_r)
abcd.data.cbcls$cbcl_scr_dsm5_opposit_r<-as.numeric(abcd.data.cbcls$cbcl_scr_dsm5_opposit_r)

summary(lmer(cbcl_scr_dsm5_opposit_r~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance+monitoring,data=abcd.data.cbcls))

### What about parental warmth x monitoring?
```{r}
summary(lmer(thompson_PC1~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance*monitoring,data=abcd.data)) # *
summary(lmer(thompson_PC2~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance*monitoring,data=abcd.data)) #
summary(lmer(thompson_PC3~interview_age+sex+(1|rel_family_id)+(1|abcd_site)+acceptance*monitoring,data=abcd.data)) # .
```
